\textsf{Are large language models superhuman chemists?}

### 2024-10-31 src/tex/appendix.tex
update formatting
Former-commit-id: 501f51ac185563ad2cbff39867d808e4817163eb
last cleanup
Former-commit-id: 1cdaf1565bd60c21aa867e8086408b4de3b56b2f

\begin{figure}[htb]
    \centering
    \includegraphics[width=\textwidth]{figures/log_probs_calibration_plot_overall_filtered.pdf}
    \caption{\textbf{Reliability diagram of histogram of logit-based confidence estimates.} For this analysis we obtained the linear probability from the logprobs of the models. Only logprobs of the tokens corresponding to the answers were considered. Linear proabability was computed by taking exponential of logprobs (for sequences with multiple tokens, values were multiplied).  The plot shows the average predicted probability and the fraction of correct answers for each bin of linear probabilities. The ideal scenario is a diagonal line, indicating perfect calibration where the model's confidence aligns perfectly with the actual correctness. The \gls{ece} value quantifies the overall calibration performance, with a lower \gls{ece} indicating better calibration.}
    \label{fig:confidence_score_distributions}
    \script{plot_logprobs.py}
\end{figure}

>>>>

\begin{figure}[htb]
    \centering
    \includegraphics[width=\textwidth]{figures/log_probs_calibration_plot_overall_filtered.pdf}
    \caption{\textbf{Reliability diagram of logit-based confidence estimates.} For this analysis we obtained the linear probability from the logprobs of the models. Only logprobs of the tokens corresponding to the answers were considered. Linear probability was computed by taking exponential of logprobs (for sequences with multiple tokens, values were multiplied).  The plot shows the average predicted probability and the fraction of correct answers for each bin of linear probabilities. The ideal scenario is a diagonal line, indicating perfect calibration where the model's confidence aligns perfectly with the actual correctness. The \gls{ece} value quantifies the overall calibration performance, with a lower \gls{ece} indicating better calibration.}
    \label{fig:confidence_score_distributions}
    \script{plot_logprobs.py}
\end{figure}

### 2024-10-31 src/tex/ms.tex
last cleanup
Former-commit-id: 1cdaf1565bd60c21aa867e8086408b4de3b56b2f

 \begin{figure}[!h]
     \centering
     \includegraphics[width=\textwidth]{figures/model_confidence_performance.pdf}
     \caption{\textbf{Calibration and distribution of confidence estimates.} For this analysis, we used verbalized confidence estimates from the model. We prompted the models to return a confidence score on an ordinal scale to obtain those estimates.
     The line plot shows the average fraction of correctly answered questions for each confidence level. The bar plot shows the distribution of confidence estimates.
     A confidence estimate would be well-calibrated if the average fraction of correctly answered questions increases with the confidence level.      The dashed black line indicates this ideal behavior, which would be monotonically increasing correctness with higher
     levels of confidence.
     We find that most models are not well-calibrated and provide misleading confidence estimates.
     }
     \label{fig:confidence_vs_performance}
     \script{confidence_estimate.py}
 \end{figure}

@@

\input{sections/human_subset_selection.tex}

>>>>

 \begin{figure}[!h]
     \centering
     \includegraphics[width=\textwidth]{figures/model_confidence_performance.pdf}
     \caption{\textbf{Reliability and distribution of confidence estimates.} For this analysis, we used verbalized confidence estimates from the model. We prompted the models to return a confidence score on an ordinal scale to obtain those estimates.
     The line plot shows the average fraction of correctly answered questions for each confidence level. The bar plot shows the distribution of confidence estimates.
     A confidence estimate would be well-calibrated if the average fraction of correctly answered questions increases with the confidence level.      The dashed black line indicates this ideal behavior, which would be monotonically increasing correctness with higher
     levels of confidence.
     We find that most models are not well-calibrated and provide misleading confidence estimates.
     }
     \label{fig:confidence_vs_performance}
     \script{confidence_estimate.py}
 \end{figure}

@@

\paragraph{Question selection} \label{sec:subset-selection}

 Several design choices were made when selecting \chembenchmini. Firstly, from the full dataset, we kept all the questions labeled as advanced. In this way, we can obtain a deeper insight into the capabilities of \glspl{llm} on advanced tasks when compared to actual chemists. Secondly, we sample a maximum of three questions across all possible combinations of categories (i.e., knowledge or reasoning) and topics (e.g., organic chemistry, physical chemistry). Thirdly, we do not include any intuition questions in this subset because the intended use of \chembenchmini is to provide a fast and fair evaluation of \glspl{llm} independent of any human baseline. In total, \variable{output/num_human_answered_questions.txt} questions have been sampled for \chembenchmini. Then, this set is divided into two subsets based on the aforementioned combinations. One of the question subsets allows tool use, and the other does not.
